Rank Recovery

### 2020-11-11 output_scripts/rank_recovery.Rmd
add glicko comparison for test data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(rstan)
library(cmdstanr)
library(compete)
```

@@

$$
f =(0.1, 0.2, 0.4, 0.7, 0.9).
$$

@@

```{r sim data from c-mmhp, cache=TRUE}
sim_model3_data <- simulateLatentMMHP(lambda0_matrix = object_matrix$lambda0_matrix,
                                      lambda1_matrix = object_matrix$lambda1_matrix,
                                      alpha_matrix = object_matrix$alpha_matrix,
                                      beta_matrix = object_matrix$beta_matrix,
                                      q1_matrix = object_matrix$q1_matrix,
                                      q2_matrix = object_matrix$q2_matrix,
                                      horizon = 200)
clean_sim_data <- cleanSimulationData(raw_data = sim_model3_data, 
                                        cut_off = cut_off, 
                                      N = length(object_par$f_vec_1))
```

```{r recover cmmhp model, eval=FALSE ,cache=TRUE}
# model3 <- stan_model(here("lib","sim_model3.stan"))
model3 <- cmdstan_model(here("lib","sim_model3.stan"))
data_list <- list(max_Nm=max(clean_sim_data$N_count),
                    N_til = length(clean_sim_data$I_fit),
                    M=sum(clean_sim_data$N_count>=cut_off),
                    N=length(object_par$f_vec_1),
                    I_fit = clean_sim_data$I_fit,
                    J_fit = clean_sim_data$J_fit,
                    T = tail(clean_sim_data$day_hour,1),
                    Nm = as.vector(clean_sim_data$N_count[clean_sim_data$N_count 
                                                          >= cut_off]),
                    event_matrix = clean_sim_data$event_matrix,
                    interevent_time_matrix = clean_sim_data$time_matrix,
                    max_interevent = clean_sim_data$max_interevent)

model3_stan_fit <- model3$sample(data = data_list,
                                seed = 123,
                                iter_warmup = 50,
                                iter_sampling = 50,
                                chains = 4, 
                                parallel_chains = 4,
                                refresh = 100)
model3_stan_fit$save_object(file = here("output", "rank_sims", 
                                        "rmd_stan",
                                        "model_3_stan_fit.RDS"))

@@

```{r extract stan fit}
model3_stan_fit <- readRDS(here("output", "rank_sims", 
                                        "rmd_stan",
                                        "model_3_stan_fit.RDS"))
stanfit <- rstan::read_stan_csv(model3_stan_fit$output_files())

stansims <- rstan::extract(stanfit)
## truth is .1, .2, .4, .7, .9

apply(stansims$f, 2, mean)
# posterior mean
# posterior sd
# apply(stansims$f, 2, sd)

# most important is the order
m3_rank <- order(apply(stansims$f, 2, mean))
#order(apply(stansims$f, 2, sd))
```
We can compare this to the ranking from I&SI on the count data alone, which
is given below.

```{r fit I&SI, results='hide'}
count_data <- get_wl_matrix(df = cbind(clean_sim_data$start,clean_sim_data$end))
isi.out <- compete::isi98(m = count_data, random = TRUE)
isi_rank <- as.numeric(rev(isi.out$best_order)) # ranked from lowest to highest
```
```{r print isi result}
isi_rank
```

@@

```{r}
true_rank <- c(1:5)

@@

```{r sim dc mmhp, cache=TRUE}
gamma_var <- c(0.01, 0.02, 0.03, 0.06, 0.07)
zeta_var <- c(0.075, 0.02, 0.03, 0.05, 0.08 )
lambda_dc_matrix <- outer(gamma_var, zeta_var, "+")
# lambda_dc_matrix

@@

@@

```{r fit dc mmhp, eval=FALSE}
model3_dc <- cmdstan_model(here("lib","sim_model3_dc.stan"))
data_list_dc <- list(max_Nm = max(clean_sim_data_dc$N_count),
                    N_til = length(clean_sim_data_dc$I_fit),
                    M=sum(clean_sim_data_dc$N_count>=cut_off),
                    N=length(object_par$f_vec_1),
                    I_fit = clean_sim_data_dc$I_fit,
                    J_fit = clean_sim_data_dc$J_fit,
                    T = tail(clean_sim_data_dc$day_hour,1),
                    Nm = as.vector(clean_sim_data_dc$N_count[
                      clean_sim_data_dc$N_count 
                                                          >= cut_off]),
                    event_matrix = clean_sim_data_dc$event_matrix,
                    interevent_time_matrix = clean_sim_data_dc$time_matrix,
                    max_interevent = clean_sim_data_dc$max_interevent)

@@

```{r load dc stan fit}
model3_dc_stan_fit <- readRDS(here("output", "rank_sims", 
                                        "rmd_stan",
                                        "model_3_dc_stan_fit.RDS"))
stanfit_dc <- rstan::read_stan_csv(model3_dc_stan_fit$output_files())
stansims_dc <- rstan::extract(stanfit_dc)
## truth is .1, .2, .4, .7, .9
apply(stansims_dc$f, 2, mean)
apply(stansims_dc$f, 2, sd)

# most important is the order
m3_dc_rank <- sort(apply(stansims_dc$f, 2, mean))
```

Then we can compare the I&SI model here also.

```{r i and si dc}
count_data_dc <- get_wl_matrix(df = cbind(clean_sim_data_dc$start, 
                                       clean_sim_data_dc$end))
isi_dc.out <- compete::isi98(m = count_data_dc, random = TRUE)
isi_rank_dc <- as.numeric(rev(isi_dc.out$best_order)) # ranked from lowest to highest

isi_rank_dc
true_rank <- c(1:5)

@@

So when we consider this degree corrected model we also
correctly recover the latent ranking order in this simulation from 
both the stan fit and the I&SI model.

@@

We also wish to consider how well we can infer the latent rank 
over time, where

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy = TRUE)
library(here)
library(tidyverse)
library(rstan)
library(cmdstanr)
library(compete)
library(posterior)
```

@@

$$
f = (0.1, 0.2, 0.4, 0.7, 0.9).
$$

@@

```{r sim data from c-mmhp, eval=FALSE}
sim_model3_data <- simulateLatentMMHP(lambda0_matrix = object_matrix$lambda0_matrix,
                                      lambda1_matrix = object_matrix$lambda1_matrix,
                                      alpha_matrix = object_matrix$alpha_matrix,
                                      beta_matrix = object_matrix$beta_matrix,
                                      q1_matrix = object_matrix$q1_matrix,
                                      q2_matrix = object_matrix$q2_matrix,
                                      horizon = 200)
clean_sim_data <- cleanSimulationData(raw_data = sim_model3_data, 
                                        cut_off = cut_off, 
                                      N = length(object_par$f_vec_1))

saveRDS(clean_sim_data,file = here("output", "rank_sims", 
                                        "rmd_stan",
                                        "sim_data_M3.RDS"))

```{r recover cmmhp model, eval=FALSE}
# model3 <- stan_model(here("lib","sim_model3.stan"))
clean_sim_data <- readRDS(file = here("output", "rank_sims", 
                                        "rmd_stan",
                                        "sim_data_M3.RDS"))
model3 <- cmdstan_model(here("lib","sim_model3.stan"))
data_list <- list(max_Nm=max(clean_sim_data$N_count),
                    N_til = length(clean_sim_data$I_fit),
                    M=sum(clean_sim_data$N_count>=cut_off),
                    N=length(object_par$f_vec_1),
                    I_fit = clean_sim_data$I_fit,
                    J_fit = clean_sim_data$J_fit,
                    T = tail(clean_sim_data$day_hour,1),
                    Nm = as.vector(clean_sim_data$N_count[clean_sim_data$N_count 
                                                          >= cut_off]),
                    event_matrix = clean_sim_data$event_matrix,
                    interevent_time_matrix = clean_sim_data$time_matrix,
                    max_interevent = clean_sim_data$max_interevent)

model3_stan_fit <- model3$sample(data = data_list,
                                seed = 123,
                                iter_warmup = 500,
                                iter_sampling = 500,
                                chains = 4, 
                                parallel_chains = 4,
                                refresh = 100)
model3_stan_fit$save_object(file = here("output", "rank_sims", 
                                        "rmd_stan",
                                        "model_3_stan_fit.RDS"))

@@

```{r extract stan fit}
model3_stan_fit <- readRDS(here("output", "rank_sims", 
                                        "rmd_stan",
                                        "model_3_stan_fit.RDS"))

m3_draws <- as_draws_df(model3_stan_fit$draws())
m3_post_rank <- m3_draws %>% select(starts_with("f")) %>% apply(2,mean)
m3_post_rank
## truth is .1, .2, .4, .7, .9

# posterior mean
# posterior sd
# apply(stansims$f, 2, sd)

# most important is the order
m3_rank <- order(m3_post_rank)
#order(apply(stansims$f, 2, sd))
```

We can compare this to the ranking from I&SI on the count data alone, which
is given below. This incorrectly orders the two lowest ranked nodes.

```{r fit I&SI, results='hide'}
clean_sim_data <- readRDS(file = here("output", "rank_sims", 
                                        "rmd_stan",
                                        "sim_data_M3.RDS") )
count_data <- get_wl_matrix(df = cbind(clean_sim_data$start,clean_sim_data$end))
isi.out <- compete::isi98(m = count_data, random = TRUE)
isi_rank <- as.numeric(rev(isi.out$best_order)) # ranked from lowest to highest
```
```{r print isi result}
isi_rank
```

@@

```{r overall sim ranks}
true_rank <- c(1:5)

@@

```{r sim dc mmhp, eval=FALSE}
gamma_var <- c(0.01, 0.02, 0.03, 0.06, 0.07)
zeta_var <- c(0.075, 0.02, 0.03, 0.05, 0.08 )
lambda_dc_matrix <- outer(gamma_var, zeta_var, "+")
# lambda_dc_matrix

@@

saveRDS(clean_sim_data_dc,file = here("output", "rank_sims", 
                                        "rmd_stan",
                                        "sim_data_M3_dc.RDS"))

@@

```{r fit dc mmhp, eval=FALSE}
clean_sim_data_dc <- readRDS(file = here("output", "rank_sims", 
                                        "rmd_stan",
                                        "sim_data_M3_dc.RDS"))
model3_dc <- cmdstan_model(here("lib","sim_model3_dc.stan"))
data_list_dc <- list(max_Nm = max(clean_sim_data_dc$N_count),
                    N_til = length(clean_sim_data_dc$I_fit),
                    M=sum(clean_sim_data_dc$N_count>=cut_off),
                    N=length(object_par$f_vec_1),
                    I_fit = clean_sim_data_dc$I_fit,
                    J_fit = clean_sim_data_dc$J_fit,
                    T = tail(clean_sim_data_dc$day_hour,1),
                    Nm = as.vector(clean_sim_data_dc$N_count[
                      clean_sim_data_dc$N_count 
                                                          >= cut_off]),
                    event_matrix = clean_sim_data_dc$event_matrix,
                    interevent_time_matrix = clean_sim_data_dc$time_matrix,
                    max_interevent = clean_sim_data_dc$max_interevent)

@@

```{r load dc stan fit}
model3_dc_stan_fit <- readRDS(here("output", "rank_sims", 
                                        "rmd_stan",
                                        "model_3_dc_stan_fit.RDS"))
m3_dc_draws <- as_draws_df(model3_dc_stan_fit$draws())
m3_dc_rank <- m3_dc_draws %>% select(starts_with("f")) %>% apply(2,mean)
m3_dc_rank
## truth is .1, .2, .4, .7, .9
# apply(stansims_dc$f, 2, sd)

# most important is the order
m3_dc_rank <- order(m3_dc_rank)
```

Then we can compare the I&SI model here also. Here there is exact agreement.

```{r i and si dc, results='hide'}
clean_sim_data_dc <- readRDS(file = here("output", "rank_sims", 
                                        "rmd_stan",
                                        "sim_data_M3_dc.RDS") ) 
count_data_dc <- get_wl_matrix(df = cbind(clean_sim_data_dc$start, 
                                       clean_sim_data_dc$end))
isi_dc.out <- compete::isi98(m = count_data_dc, random = TRUE)
isi_rank_dc <- as.numeric(rev(isi_dc.out$best_order)) 
# ranked from lowest to highest
```

```{r tables of agreement}
isi_rank_dc
true_rank <- c(1:5)

@@

So when we consider this degree corrected model we
correctly recover the latent ranking order in this simulation from 
both the stan fit and the I&SI model.

## Multiple Simulations

To confirm this result we simulate and fit these models to data from 
the degree corrected model 20 times. We wish to see how many times 
they agree.

We see that the C-MMHP model somewhat better recovers the true
ranks in this simulation, with both performing well. We also
see that any discrepancies from the true ranking are small, with the
order of a pair of nodes being swapped.

```{r load in sims, options}
sim_files <- list.files(here("output","rank_sims","sims_m3_dc_isi"))

all_sims <- tibble()
for(i in seq_along(sim_files)) {
 single_fit <- readRDS(here("output", "rank_sims", 
                            "sims_m3_dc_isi",
                            sim_files[i]))
 all_sims <- all_sims %>% bind_rows(single_fit)
}

We can look at the overall summary of the ranking of the two models and
the truth.

```{r summary ranks}
all_sims %>% 
  group_by(truth,m3_dc,isi) %>%
  count() %>% arrange(-n) 
```

We would like to identify how many times each model correctly recovers the
full ranking.
We see that the C-MMHP is fully correct 80% of the time while the 
I&SI ranking is fully correct in 60% of these simulations.

```{r prop full recover}
all_sims$sim <- rep(1:20, each = 5)

all_sims %>% select(sim, truth, m3_dc) %>%
  mutate(diff = truth - m3_dc) %>%
  group_by(sim) %>%
  summarise(prop_diff = mean(diff!=0)) %>%
  summarise(full_recovery = mean(prop_diff == 0))

all_sims %>% select(sim, truth, isi) %>%
  mutate(diff = truth - isi) %>%
  group_by(sim) %>%
  summarise(prop_diff = mean(diff!=0)) %>%
  summarise(full_recovery = mean(prop_diff == 0))
```

We see strong agreement between our model and the true ranking of
the simulated data for most nodes.

```{r truth_m3}
all_sims %>%
  group_by(truth, m3_dc) %>%
  count() %>%
  mutate(per = n/20) %>%
  arrange(-per)
```

Finally, we look at the agreement between the ISI ranking and the truth.

```{r truth isi}
all_sims %>%
  group_by(truth, isi) %>%
  count() %>%
  mutate(per = n/20) %>%
  arrange(-per)

@@

We also wish to consider how well we can infer the latent rank 
over time. To investigate this, we wish to split the data in a Mice cohort
(or simulation data) temporally, and examine the inferred model on the 
first half of the interactions compared to on the remaining data,
by looking at the inferred I&SI scores.

```{r compute isi}
# taking code from compete package
compute_isi <- function(win_loss_matrix, curr_ranking) {
  # this takes ranking from highest to lowest
  wl <- win_loss_matrix[curr_ranking,curr_ranking]
  n <- as.numeric(dim(wl)[1])
  m <- matrix(0, n, n)
  siind <- matrix(0, n, n)
  I = 0
  SI = 0
  for (j in 2:n) {
      for (i in 1:(j - 1)) {
          siind[j, i] = j - i
          siind[i, j] = j - i
          if (wl[j, i] == 0 & wl[i, j] == 0) {
              m[j, i] = 0
              m[i, j] = 0
          }
          else {
              if (wl[j, i] > wl[i, j]) {
                m[j, i] = 1
                m[i, j] = -1
              }
              if (wl[j, i] < wl[i, j]) {
                m[j, i] = -1
                m[i, j] = 1
              }
              if (wl[j, i] == wl[i, j]) {
                m[j, i] = 0.5
                m[i, j] = 0.5
              }
          }
      }
  }
  I = sum(lower.tri(m) * m == 1)
  SI = sum((lower.tri(m) * m == 1) * siind)

  list(I_score = I, SI_score = SI, wl_mat = wl)
}

```{r check isi function, results='hide'}
clean_sim_data <- readRDS(file = here("output", "rank_sims", 
                                        "rmd_stan",
                                        "sim_data_M3.RDS"))
clean_sim_data_dc$N_count

count_data_dc <- get_wl_matrix(df = cbind(clean_sim_data_dc$start, 
                                       clean_sim_data_dc$end))
isi_dc.out <- compete::isi98(m = count_data_dc, random = TRUE)

# these agree

compute_isi(count_data_dc, curr_ranking = c(5,3,1,2,4))

We want to fit our models on the first half of the data
and then compare the ranking estimated from it on 
the remaining data, in terms of I&SI score.

```{r sim data and split, eval=FALSE}
sim_model3_data_dc <- 
  simulateLatentMMHP(lambda0_matrix = object_dc_matrix$lambda0_matrix,
                      lambda1_matrix = object_dc_matrix$lambda1_matrix,
                      alpha_matrix = object_dc_matrix$alpha_matrix,
                      beta_matrix = object_dc_matrix$beta_matrix,
                      q1_matrix = object_dc_matrix$q1_matrix,
                      q2_matrix = object_dc_matrix$q2_matrix,
                      horizon = 200)

sim_data <- tibble(start = sim_model3_data_dc$start,
                   end = sim_model3_data_dc$end,
                   day_hour = sim_model3_data_dc$day_hour)

training_data <- sim_data %>% filter(day_hour <= 100)
test_data <- sim_data %>% filter(day_hour > 100 )

saveRDS(training_data,file = here("output", "rank_sims", 
                                        "rmd_stan",
                                        "train_data_dc.RDS"))
saveRDS(test_data,file = here("output", "rank_sims", 
                                        "rmd_stan",
                                        "test_data_dc.RDS"))

```{r fit m3 to first half, eval=FALSE}
training_data <- readRDS(file = here("output", "rank_sims", 
                                        "rmd_stan",
                                        "train_data_dc.RDS"))

clean_training <- cleanSimulationData(training_data, cut_off = 3, N = 5)

data_list_training <- list(max_Nm = max(clean_training$N_count),
                    N_til = length(clean_training$I_fit),
                    M=sum(clean_training$N_count>=cut_off),
                    N=length(object_par$f_vec_1),
                    I_fit = clean_training$I_fit,
                    J_fit = clean_training$J_fit,
                    T = tail(clean_training$day_hour,1),
                    Nm = as.vector(clean_training$N_count[
                      clean_training$N_count 
                                                          >= cut_off]),
                    event_matrix = clean_training$event_matrix,
                    interevent_time_matrix = clean_training$time_matrix,
                    max_interevent = clean_training$max_interevent)

model3_dc <- cmdstan_model(here("lib","sim_model3_dc.stan"))

model3_dc_train <- model3_dc$sample(data = data_list_training,
                                seed = 123,
                                iter_warmup = 500,
                                iter_sampling = 500,
                                chains = 4, 
                                parallel_chains = 4,
                                refresh = 100)

model3_dc_train$save_object(file = here("output", "rank_sims",
                                        "rmd_stan",
                                        "model_3_dc_train_fit.RDS"))

```{r get rank from this,results='hide'}
model3_dc_train <- readRDS(here("output", "rank_sims",
                                        "rmd_stan",
                                        "model_3_dc_train_fit.RDS"))

m3_draws_train <- as_draws_df(model3_dc_train$draws())
m3_train_rank <- m3_draws_train %>% select(starts_with("f")) %>% apply(2,mean)

m3_train_rank

m3_train_rank <- rev(order(m3_train_rank))
# then feed this ranking into the test data

test_data <- readRDS(file = here("output", "rank_sims", 
                                        "rmd_stan",
                                        "test_data_dc.RDS"))

We can compare the I&SI score from the posterior ranks from the
first half of the data with that obtained from using the I&SI
ranks from the first half of the data, which agree here.

```{r fit isi train, results='hide'}

test_wl <- get_wl_matrix(cbind(test_data$start,test_data$end))
m3_test_score <- compute_isi(test_wl,curr_ranking = m3_train_rank)

# then compare this to isi for test data

training_data <- readRDS(file = here("output", "rank_sims", 
                                        "rmd_stan",
                                        "train_data_dc.RDS"))

train_wl <- get_wl_matrix(cbind(training_data$start,training_data$end))

isi_train <- isi98(train_wl)
isi_train_rank <- isi_train$best_order

isi_test_score <- compute_isi(test_wl,curr_ranking = isi_train_rank)

```{r}
m3_test_score
isi_test_score
```

How does this compare to the optimal score using only the test data.
```{r fit isi to test}
isi98(test_wl)
```

## Aggregate Ranking Model

We can similarly do this for some of the other models we use
as comparison, such as the Aggregate ranking model.

```{r fit agg rank training, results='hide'}
training_data <- readRDS(file = here("output", "rank_sims", 
                                        "rmd_stan",
                                        "train_data_dc.RDS"))

clean_training <- cleanSimulationData(training_data, cut_off = 3, N = 5)

agg_rank_training <- clean_training$N_count

agg_rank_model <- cmdstan_model(here("lib","latent_rank_agg_sim.stan"))

agg_rank_fit <- agg_rank_model$sample(data = list(n = 5,
                                                  n_matrix = agg_rank_training),
                                      seed = 123,
                                      iter_warmup = 500,
                                      iter_sampling = 500,
                                      chains = 4, 
                                      parallel_chains = 4,
                                      refresh = 100)

agg_rank_draws_train <- as_draws_df(agg_rank_fit$draws())
agg_train_mean_rank <- agg_rank_draws_train %>%
  select(starts_with("x")) %>% apply(2,mean)

agg_rank_train <- order(agg_train_mean_rank)
```

```{r show agg rank mean}
agg_train_mean_rank
```

The aggregate ranking model struggles to correctly identify the correct
ranking, resulting in a large I&SI score.

```{r compute_isi for agg rank}
test_data <- readRDS(file = here("output", "rank_sims", 
                                        "rmd_stan",
                                        "test_data_dc.RDS"))
test_wl <- get_wl_matrix(cbind(test_data$start, test_data$end))

agg_test_score <- compute_isi(test_wl,curr_ranking = agg_rank_train)
agg_test_score
```

## Glicko Model

We can also perform this comparison with the Glicko 
model on the training data.

```{r fit glicko}
library(PlayerRatings)
source(here("lib","myGlicko.R"))

training_data <- readRDS(file = here("output", "rank_sims", 
                                        "rmd_stan",
                                        "train_data_dc.RDS"))

training_data
### need score (1) for all of these, along with the row number

glick_df <- training_data %>% 
  mutate(id = row_number(), score = 1) %>%
  select(id, start, end , score)

gl_train <- my_glicko(glick_df, history=TRUE, cval=2)

gl_train

gl_ranks <- rev(gl_train$ratings$Player)
# highest to lowest

This correctly recovers the true ranking and will therefore result in 
the same optimal I&SI score.

```{r glicko test}
test_data <- readRDS(file = here("output", "rank_sims", 
                                        "rmd_stan",
                                        "test_data_dc.RDS"))
test_wl <- get_wl_matrix(cbind(test_data$start, test_data$end))

glicko_test_score <- compute_isi(test_wl,curr_ranking = gl_ranks)
glicko_test_score
